Selecting and ranking interesting pairs of cyclic temporal granularities

### 2020-11-19 main.Rmd
prelude to normalisation built

### Need for normalisation

Median and maximum distances are affected by the number of categories considered and hence the distance measure $MMPD_{raw}$, which is a combination of median and maximum would also be influenced by the number of levels. It would have higher values if $C_i$ or $C_j$ has higher levels. We would ideally want a higher value of the measure if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. In this example, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 3, 4, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which MMPD_raw has been constructed 100 times to compute the distribution. From the histogram we could see that both the shape and scale of the distribution are different when we move across different levels. More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.

This implies we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure, so that we could correctly compare $MMPD_{raw}$ across different combinations of facet and x-axis levels.

@@

```{r hist_plot, fig.cap="density mmpd raw"}
 hist_plot = mmpd_raw %>% 
      ggplot() +
      geom_density(aes(x = raw_mmpd), fill = "blue") +
      facet_grid(nx~nfacet)
hist_plot
```

>>>>

### Distribution of $MMPD_{raw}$
<!-- ### Need for normalisation -->

From Figure \ref{fig:dist-across-facets} and \ref{fig:dist-across-x} shows the MMPD distribution is different for different levels of facets and x-axis levels. With increasing number of facets, the location of the distribution shift rightwards and with increasing x-axis levels, the scale of the distribution reduces.

Median and maximum distances are affected by the number of categories considered and hence the distance measure $MMPD_{raw}$, which is a combination of median and maximum would also be influenced by the number of levels. It would have higher values if $C_i$ or $C_j$ has higher levels. We would ideally want a higher value of the measure if there is significant difference between distributions across facet or x-axis categories, and not because the number of categories are higher. In Figures \ref{fig:dist-across-facets} and \ref{fig:dist-across-x}, both the cyclic granularities $C_i$ and $C_j$ are considered such that their levels vary in the range $(2, 5,7 ,9, 14, 21, 30, 45)$. Each of these combinations is considered a panel for each of which $MMPD_{raw}$ has been constructed $100$ times to compute the distribution. <!--More specifically, Figure \ref{fig:mean-sd-max} shows the marginal contribution of increasing levels of $C_i$ and $C_j$ on distribution of $MMPD_raw$. This shows that with the increase in the number of levels of $C_i$, the location of the distribution shift rightward, whereas for increasing x-axis levels, the scale of the distribution decreases.-->

Therefore, in order to compare $MMPD_{raw}$ across different combinations of facet and x-axis levels, we need to eliminate the impact of different levels of the facets and x-axis first and get a normalized measure.

```{r dist-across-facets, fig.cap = "Distribution of $MMPD_{raw}$ for different facet levels are shown. The location of the distribution shifts rightwards implying that as the number of facets increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained. "}
mmpd_raw = read_rds("../hakear-drake/data/data_all23457914n.rds") 
mmpd_xmore = read_rds("../hakear-drake/data/data_all213045x.rds") 
mmpd_fmore = read_rds("../hakear-drake/data/data_all_213045f_allx.rds")

data_all = bind_rows(mmpd_raw, mmpd_xmore)

data_long = data_all %>% pivot_longer(c(1:2), names_to = "category", values_to = "levels")

g1 <- data_all %>% 
  filter(!(nx %in% c(3, 4))) %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(x = raw_mmpd, 
                                    y = as.factor(nfacet))) +
  facet_wrap(~nx, nrow = 2)
g1
```

```{r dist-across-x, fig.cap = "Distribution of $MMPD_{raw}$ for different x levels are shown. The scale of the distribution becomes less and less implying that as the number of x levels increase, it is more likely that a higher value of $MMPD_{raw}$ will be obtained as the spread of the distribution decreases. "}
g2 <- data_all %>% 
  filter(nfacet<=14, nfacet!=3) %>% 
  ggplot() +
  geom_boxplot(aes(y = raw_mmpd, 
                                    x = as.factor(nx))) +
  facet_wrap(~nfacet) + coord_flip()

#  hist_plot = data_all %>% 
#       ggplot() +
#       geom_density(aes(x = raw_mmpd), fill = "blue") +
#       facet_grid(nx~nfacet)
# hist_plot
g2
```

@@
